---
title: "Clustering"
author: "Alexis Bryant"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(cluster)
  library(parallel)
  library(viridis)
  install.packages("remotes")
  install.packages("ggrepel")
  library(remotes)
  library(ggrepel)

})
```
```{r}
remotes::install_version("plotly", version =  "4.11.0")
library(plotly)
install.packages("rtools")

```


## Task 1

Define Data Generator
```{r}
generate_hypercube_clusters <- function(n, k, side_length, noise_sd = 1.0) {
  centers <- diag(rep(side_length, n))  
  data_list <- lapply(1:n, function(i) {
    tibble(matrix(rnorm(k * n, mean = centers[i, ], sd = noise_sd), ncol = n))
  })
  data <- bind_rows(data_list)
  names(data) <- paste0("x", 1:n)
  data
}
```


Function that computes estimated number of clusters
```{r}
estimate_k <- function(n, side_length) {
  data <- generate_hypercube_clusters(n, 100, side_length)
  gap <- clusGap(
    data,
    FUN = function(x, k) kmeans(x, centers = k, nstart = 20, iter.max = 50),
    K.max = n + 2,
    B = 10
  )
  which.max(gap$Tab[, "gap"])
}
```

Run Simulation
```{r}
params <- expand_grid(
  n = c(6, 5, 4, 3, 2),
  side_length = seq(10, 1, by = -1)
)
estimate_k_safe <- safely(estimate_k, otherwise = NA_real_)

results_fast <- params %>%
  mutate(estimated_k = map2(n, side_length, ~ estimate_k_safe(.x, .y))) %>%
  mutate(
    estimated_k = map_dbl(estimated_k, function(x) {
      if (is.null(x$result) || length(x$result) == 0) NA_real_ else x$result
    })
  )
```

Visualize Results
```{r}
p1 <- ggplot(results_fast, aes(side_length, estimated_k, color=factor(n))) +
geom_line(linewidth=1.2) + geom_point(size=3) +
geom_hline(aes(yintercept=n), linetype="dashed", color="black") +
scale_x_reverse() + scale_color_viridis_d() +
labs(title="Gap Statistic: Hypercube Cluster", x="Side Length", y="Estimated k", color="Dim (n)") +
theme_minimal(base_size=14)
p1

```



### At what point do distinct groups of data become so close that our algorithms can no longer tell them apart? For each dimension, at what side_length does the gap statistic method consistently begin to reduce its estimate of the number of clusters?


As side_length decreases the clusters become harder to distinguish. Meaning that the gap statistic underestimates k once the clusters begin to overlap.The side length were the estimated k is greater than n marks the failure point for that specific dimension which according to the visualizations would be around 1.25-2.5. 

Dimensions 2,3, and 6 start decreasing at about length 3
Dimension 4 starts decreasing at about length 2
Dimension 5 starts increasing at about length 4

## Task 2 Spectral Clustering on Concentric Shells

Generate Shell Data
```{r}
generate_shell_clusters <- function(n_shells, k_per_shell, max_radius, noise_sd = 0.1) {
radii <- seq(1, max_radius, length.out = n_shells)
shells <- lapply(radii, function(r) {
theta <- runif(k_per_shell, 0, 2*pi)
phi <- acos(runif(k_per_shell, -1, 1))
x <- (r + rnorm(k_per_shell, 0, noise_sd)) * sin(phi) * cos(theta)
y <- (r + rnorm(k_per_shell, 0, noise_sd)) * sin(phi) * sin(theta)
z <- (r + rnorm(k_per_shell, 0, noise_sd)) * cos(phi)
tibble(x, y, z, shell = as.factor(r))
})
bind_rows(shells)
}

```

Spectral Clustering Wrapper
```{r}
spectral_clustering <- function(x, k, d_threshold = 1) {
  x <- as.matrix(x)
  dist_mat <- as.matrix(dist(x))
  A <- (dist_mat < d_threshold) * 1
  diag(A) <- 0
  D <- diag(rowSums(A))

  # ---- guard against empty / isolated nodes ----
  zero_deg <- which(diag(D) == 0)
  if (length(zero_deg) > 0) {
    # connect isolated points weakly to themselves
    diag(D)[zero_deg] <- 1e-6
  }

  L <- D - A
  D_inv_sqrt <- diag(1 / sqrt(diag(D)))
  L_sym <- D_inv_sqrt %*% L %*% D_inv_sqrt

  # Replace any remaining NaN / Inf with 0
  L_sym[!is.finite(L_sym)] <- 0

  eig <- eigen(L_sym, symmetric = TRUE)
  if (ncol(eig$vectors) < k) k <- ncol(eig$vectors)
  U <- eig$vectors[, (ncol(eig$vectors) - k + 1):ncol(eig$vectors)]
  kmeans(U, centers = k, nstart = 10)
}

```

Gap Stat
```{r}
radii <- seq(10, 0, by=-1)
n_shells <- 4
k_per_shell <- 100

results_spec <- tibble(max_radius = radii) %>%
mutate(estimated_k = map_dbl(max_radius, function(r) {
data <- generate_shell_clusters(n_shells, k_per_shell, r)
gap <- clusGap(data[, c("x","y","z")],
FUN = function(x, k) spectral_clustering(x, k, d_threshold=1),
K.max = 6, B = 10)
which.max(gap$Tab[, "gap"])
}))

```

Visualize Results
```{r}
ggplot(results_spec, aes(max_radius, estimated_k)) +
geom_line(color="steelblue", linewidth=1.2) +
geom_point(size=3, color="darkred") +
geom_hline(yintercept=4, linetype="dashed", color="black") +
scale_x_reverse() +
labs(title="Spectral Clustering on Concentric Shells",
x="Maximum Radius", y="Estimated Number of Clusters") +
theme_minimal(base_size=14)
```


### Explain the algorithm’s failure point in your plot as a function of the decreasing max_radius, and then discuss how this result would change if you were to use a different d_threshold.


In this simulation spectral clustering identifies four clusters when the shells are well separated, but fails once the maximum radius drops below approximately 3–4. At that point, the adjacency matrix becomes dense enough that separate shells merge into one connected region. Adjusting the neighborhood (d_threshold) alters this behavior: smaller thresholds preserve separability longer by limiting cross-cluster edges, while larger thresholds accelerate failure by increasing inter-shell connectivity.So, the algorithm’s effectiveness depends critically on both the geometric separation of clusters and the chosen scale of connectivity.
