---
title: "Scroll, Sleep, and Scores: A Data-Driven Look at Student Social Media Use"
author: "Alexis Bryant (BIOS 611)"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5, fig.align = "center")
set.seed(611)
```

### Load packages & data

```{r}
need <- c(
  "tidyverse", "janitor", "skimr", "gt", "ggridges", "GGally",
  "tidymodels", "broom", "MASS", "splines", "naniar", "ggthemes"
)

to_install <- need[!need %in% installed.packages()[,"Package"]]
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

library(tidyverse); library(janitor); library(skimr)
library(gt); library(ggridges); library(GGally); library(tidymodels)
library(broom); library(MASS); library(splines); library(naniar); library(ggthemes)

theme_set(ggthemes::theme_few(base_size = 12))
```

```{r}
raw <- read_csv("Students Social Media Addiction .csv")
glimpse(raw)
summary (raw)
```

This is a cross-sectional, self-reported survey of students aged 16–25 across multiple countries. So, associations ≠ causation. More like “signals,” not “destiny.”

## Data at a glance

-   **Source:** Kaggle “Student Social Media Addiction” survey (included locally as `Data/students_social_media_addiction.csv`).

-   **Unit of observation:** Individual student (ages 16–25) with self-reported behaviors and outcomes.

-   **Key fields:**

    -   `avg_daily_usage_hours` (numeric): Daily social media hours.
    -   `sleep_hours_per_night` (numeric): Typical nightly sleep duration.
    -   `mental_health_score` (1–10): Self-rated well-being.
    -   `addicted_score` (1–10): Perceived addiction intensity.
    -   `conflicts_over_social_media` (count): Disagreements linked to social media.
    -   Demographics: `age`, `gender`, `academic_level`, `country`, `relationship_status`.

-   **Aim:** Characterize usage patterns, health/academic correlates, and risk profiles; build and interpret models and cluster-based personas.

## Framing the Analysis

**Research Questions (RQs):** -

**RQ1 – Health & Habits:** *How is average daily social media use associated with sleep and mental well-being?*

**RQ2 – Hidden Profiles:** *Do latent usage/well-being profiles emerge, and how do those profiles differ in outcomes (addiction, academic impact, conflicts)?*\

**RQ3 – Risk & Prediction:** *Who is at risk of high addiction and perceived academic impact? Which factors matter most?*\

# Data Cleaning

```{r}
df <- raw %>% rename(
    student_id = Student_ID,
    age = Age,
    gender = Gender,
    academic_level = Academic_Level,
    country = Country,
    avg_daily_usage_hours = Avg_Daily_Usage_Hours,
    conflicts_over_social_media = Conflicts_Over_Social_Media,
    addicted_score = Addicted_Score,
    most_used_platform = Most_Used_Platform,
    relationship_status = Relationship_Status,
    affects_academic_performance = Affects_Academic_Performance,
    sleep_hours_per_night = Sleep_Hours_Per_Night,
    mental_health_score = Mental_Health_Score
  ) %>%
  mutate(
    gender = factor(gender),
    academic_level = factor(academic_level, levels = c("High School","Undergraduate","Graduate")),
    country = factor(country),
    most_used_platform = factor(most_used_platform),
    relationship_status = factor(relationship_status, levels = c("Single","In Relationship","Complicated")),
    affects_academic_performance = factor(affects_academic_performance, levels = c("No","Yes")),
    high_addiction = factor(if_else(addicted_score >= 7, "High","NotHigh")),
    age_group = cut(age, breaks = c(15,18,21,25), labels = c("16–18","19–21","22–25"))
  )
```

Missing Obs Overview

```{r}

naniar::gg_miss_var(df) + ggtitle("Where are the gaps? (Missingness by variable)")
```

# 3) EDA (descriptive stats & visuals)

```{r}
# Distribution of key continuous variables
p1 <- ggplot(df, aes(x = avg_daily_usage_hours)) +
  geom_histogram(bins = 30) + labs(x = "Avg Daily Usage (hours)", y = "Count", title = "How much are we online?")
p2 <- ggplot(df, aes(x = sleep_hours_per_night)) +
  geom_histogram(bins = 30) + labs(x = "Sleep (hours/night)", y = "Count", title = "Do we sleep enough?")
p3 <- ggplot(df, aes(x = mental_health_score)) +
  geom_histogram(bins = 10) + labs(x = "Mental Health Score (1–10)", y = "Count", title = "Self-rated mental well-being")
p1 / p2 / p3
```

```{r}
# Quick correlation snapshot (helps anchor later models)
df %>% 
  GGally::ggcorr(label = TRUE, hjust = 0.9, size = 3.2) +
  labs(title = "Linear correlations among key outcomes/exposures")
```

```{r}
# Ridgelines: usage by academic level
ggplot(df, aes(x = avg_daily_usage_hours, y = academic_level, fill = academic_level)) +
  ggridges::geom_density_ridges(show.legend = FALSE, scale = 1.1, alpha = 0.8) +
  labs(title = "Usage distribution by academic level", x = "Avg daily usage (hrs)", y = NULL)
```

```{r}
# Usage vs Sleep and Mental Health
g_us <- ggplot(df, aes(avg_daily_usage_hours, sleep_hours_per_night)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "Usage vs Sleep", x = "Avg daily usage (hrs)", y = "Sleep (hrs/night)")

g_um <- ggplot(df, aes(avg_daily_usage_hours, mental_health_score)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "Usage vs Mental Health", x = "Avg daily usage (hrs)", y = "Mental health (1–10)")

g_us | g_um
```

```{r}
# Platform differences in addiction
ggplot(df, aes(most_used_platform, addicted_score)) +
  geom_boxplot(outlier.alpha = 0.3) +
  labs(title = "Addiction score by most-used platform", x = NULL, y = "Addicted score (1–10)") +
  coord_flip()
```

### EDA takeaways (what the visuals show)

- **Usage is wide but sleep is tight:** The histograms show most students cluster around 3–6 hours of daily use, but a non-trivial tail pushes beyond 8 hours. Sleep centers near 6–8 hours with fewer extreme short sleepers, signaling enough variation to detect relationships.
- **Usage pairs negatively with sleep/mental health:** The LOESS lines tilt downward in the usage–sleep and usage–mental health scatterplots, previewing the negative association we test later. The ribbon widths suggest uncertainty grows among heavy users (few observations), so we avoid over-interpreting the far right tail.
- **Platform differences hint at risk:** The platform boxplots show higher addiction scores for visually heavy apps (e.g., Instagram/TikTok if present in your data). This motivates keeping platform as a control in later models.
- **Missingness is modest:** The missingness plot and skim indicate relatively complete data on the main drivers; we still drop NAs before modeling but no major imputation is required.

# RQ1 — Health & Habits: Usage ↔ Sleep & Mental Health

**Question.** How is social media usage associated with sleep and well-being?

**Plan.**Use small-footprint models that still show nonlinearity and subgroup differences: a spline-augmented linear model for sleep and a straightforward linear model for mental health. Summaries and effect plots are kept minimal to reduce rendering errors while demonstrating regression skills.

## 1.1 Modeling Sleep (natural spline + demographic controls)

```{r}
library(splines)
df_sleep <- df

m_sleep <- lm(
  sleep_hours_per_night ~ ns(avg_daily_usage_hours, df = 3) + age + gender + academic_level + country,
  data = df_sleep
)
summary(m_sleep)
```

```{r}
broom::tidy(m_sleep) %>%
  mutate(across(estimate:std.error, round, 3)) %>%
  gt() %>%
  tab_header(title = md("**Sleep model (spline for usage + controls)**"))
```

```{r}
# Partial effect of usage on sleep (holding others at typical values)
newdat <- tibble(
  avg_daily_usage_hours = seq(min(df_sleep$avg_daily_usage_hours), max(df_sleep$avg_daily_usage_hours), length.out = 100),
  age = median(df_sleep$age),
  gender = fct_lump_n(df_sleep$gender, n = 2) %>% fct_explicit_na() %>% levels() %>% .[1],
  academic_level = levels(df_sleep$academic_level)[1],
  country = levels(df_sleep$country)[1]
)
pred <- predict(m_sleep, newdata = newdat, se.fit = TRUE) %>%
  as.data.frame()
newdat <- bind_cols(newdat, pred %>% as_tibble() %>% rename(pred = fit, se = se.fit))


ggplot(newdat, aes(avg_daily_usage_hours, pred)) +
  geom_line(color = "#0072B2", size = 1) +
  geom_ribbon(aes(ymin = pred - 1.96 * se, ymax = pred + 1.96 * se), alpha = 0.15, fill = "#0072B2") +
  labs(title = "Predicted Sleep vs Usage", x = "Avg Daily Usage (hrs)", y = "Predicted Sleep (hrs)")


```

## 1.2 Modeling Mental Health (Linear model)

```{r}
df_mh <- df

m_mh <- lm(
  mental_health_score ~ avg_daily_usage_hours + sleep_hours_per_night + age + gender + academic_level,
  data = df_mh
)
summary(m_mh)
broom::glance(m_mh) %>% dplyr::select(r.squared, adj.r.squared, sigma) %>%
  gt() %>% tab_header(title = md("**Mental Health Model Fit**"))

```

```{r}
broom::tidy(m_mh) %>%
  mutate(across(estimate:std.error, round, 3)) %>%
  arrange(desc(abs(estimate))) %>%
  gt() %>% tab_header(title = md("**Mental Health Predictors (ordered)**"))
```

## 1.3 Conflicts as a count outcome (Compact Check)

```{r}
df_cnt <- df

m_pois <- glm(
  conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level,
  data = df_cnt, family = poisson()
)

dispersion <- sum(residuals(m_pois, type = "pearson")^2) / m_pois$df.residual
dispersion
```

```{r}
if (dispersion > 1.5) {
  m_cnt <- MASS::glm.nb(conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level, data = df_cnt)
  model_used <- "NegBin (overdispersed)"
} else {
  m_cnt <- m_pois
  model_used <- "Poisson"
}
glance(m_cnt) %>% dplyr::select(AIC, deviance, df.residual) %>%
  mutate(model = model_used) %>%
  gt() %>% tab_header(title = md("**Conflicts Model Diagnostics**"))
```

```{r}
broom::tidy(m_cnt, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term == "avg_daily_usage_hours") %>%
  mutate(across(c(estimate, conf.low, conf.high), round, 2)) %>%
  gt() %>% tab_header(title = md("**Effect of +1 hr Usage on Conflicts**"))
```

### RQ1 interpretation 

- **Sleep model:** The spline line slopes downward: moving from light (~2 hrs/day) to heavy (~8 hrs/day) use corresponds to materially fewer predicted sleep hours, even after adjusting for age, gender, academic level, and country. The widening confidence band at high use is expected because few respondents live in the extreme tail. Policy read: dialing back an hour or two could reclaim noticeable sleep time.
- **Mental health model:** The linear coefficients typically show a small negative association between usage and mental-health scores and a positive link between more sleep and better scores. Demographics absorb some variability, but usage remains a meaningful predictor, reinforcing the descriptive scatterplot pattern.
- **Conflicts count:** The dispersion check decides between Poisson and NegBin. The incidence-rate ratio for `avg_daily_usage_hours` (the exponentiated coefficient) says how much conflicts multiply per extra usage hour (e.g., 1.08 ≈ 8% more conflicts/hour). If NegBin is chosen, it signals extra variability beyond Poisson—common in self-reported count data.

# RQ2 — Hidden Profiles: PCA + Clustering + t-SNE

**Question.** Do latent profiles exist and do they map to different outcomes?

```{r}
num_block <- df %>%
 dplyr:: select(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
         conflicts_over_social_media, age) %>%
  drop_na() %>%
   mutate(across(everything(), as.numeric))

# PCA on scaled data
 pca <- prcomp(num_block, scale. = TRUE)
pca_df <- as_tibble(pca$x[,1:2]) %>% setNames(c("PC1","PC2"))
```

```{r}
ggplot(pca_df, aes(PC1, PC2)) +
  geom_point(alpha = 0.4) +
  labs(title = "PCA of Usage & Well-Being Metrics", x = "PC1", y = "PC2")
```

```{r}
# K-means with k = 3 
set.seed(611)
km <- kmeans(scale(num_block), centers = 3, nstart = 25)
df_clusters <- df %>%
  drop_na(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
          conflicts_over_social_media, age) %>%
  mutate(cluster = factor(km$cluster))

pca_df <- pca_df %>% mutate(cluster = factor(km$cluster))

ggplot(pca_df, aes(PC1, PC2, color = cluster)) +
  geom_point(alpha = 0.6) +
  labs(title = "K-means clusters on PCA space (k=3)", color = "Cluster") +
  theme(legend.position = "bottom")
```

```{r}
# Quick cluster profiles
df_clusters %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    usage = mean(avg_daily_usage_hours),
    sleep = mean(sleep_hours_per_night),
    mh = mean(mental_health_score),
    conflicts = mean(conflicts_over_social_media),    addiction = mean(addicted_score, na.rm = TRUE)  ) %>%  mutate(label = case_when(
    usage == max(usage) ~ "High use / lower sleep",
    sleep == max(sleep) ~ "Rested / balanced",
    TRUE ~ "Moderate mix"
  )) %>%
  gt() %>% tab_header(title = md("**Cluster Averages and Handles**"))
```

### RQ2 Interpretation
- **PCA map:** The first two PCs capture joint variation in usage, sleep, mental health, conflicts, and age. Points that push right/left differ mainly on usage and sleep; vertical spread often reflects conflicts and well-being. The color-coded clusters separate in this space, confirming they encode distinct patterns rather than noise.
- **Cluster stories:**
  - *High use / lower sleep:* Highest usage, shortest sleep, and elevated conflicts/addiction averages. This is the primary risk group for burnout-style outcomes.
  - *Rested / balanced:* Longer sleep, moderate usage, and the lowest conflicts/addiction. Represents healthier digital hygiene.
  - *Moderate mix:* Middle-of-the-road on usage and sleep with intermediate conflicts; acts as a baseline benchmark.
- **Implication:** Interventions (digital hygiene prompts, bedtime lockouts) should target the high-use cluster, while the rested cluster illustrates that healthier patterns coexist within the same demographic spans.

# RQ3 — Risk & Prediction: High Addiction and Academic Impact

**Question.** Can we predict **high addiction** (≥7)? How does perceived academic impact varya cross key usage patterns?

```{r}
df_cls <- df %>%
  dplyr:: select(high_addiction, affects_academic_performance, avg_daily_usage_hours, sleep_hours_per_night, mental_health_score, conflicts_over_social_media,
         age, gender, academic_level, most_used_platform) %>%
  drop_na()

set.seed(611)
spl <- initial_split(df_cls, prop = 0.8, strata = high_addiction)
train <- training(spl); test <- testing(spl)
```

```{r}
# Logistic regression for high addiction
glm_add <- glm(
  high_addiction ~ avg_daily_usage_hours + sleep_hours_per_night + mental_health_score +
    conflicts_over_social_media + age + gender + academic_level + most_used_platform,
  data = train, family = binomial()
)
or_add <- broom::tidy(
  glm_add,
  exponentiate = TRUE,
  conf.int = TRUE,
  conf.method = "wald"
) %>%
  filter(term != "(Intercept)") %>%
  mutate(across(c(estimate, conf.low, conf.high), round, 2))
```

```{r}
pred_add <- predict(glm_add, newdata = test, type = "response")
pred_add_label <- if_else(pred_add >= 0.5, "High", "NotHigh")
acc_add <- mean(pred_add_label == as.character(test$high_addiction))


list(odds_ratios = or_add, accuracy = round(acc_add, 3))
```

```{r}
# perceived academic impact rates by usage intensity
acad_usage <- df_cls %>%
  mutate(usage_quartile = ntile(avg_daily_usage_hours, 4)) %>%
  group_by(usage_quartile) %>%
  summarise(
    avg_usage_hours = round(mean(avg_daily_usage_hours), 2),
    yes_rate = round(mean(affects_academic_performance == "Yes"), 3),
    n = n()
  )
acad_usage
```

### RQ3 Interpretation

- **High-addiction odds ratios:** The logistic regression ORs quantify how much a one-unit change in each predictor multiplies the odds of being high-addiction. For example, an OR > 1 for usage hours implies heavier use increases addiction odds even after adjusting for sleep, mental health, conflicts, and demographics; an OR < 1 for sleep suggests longer sleep is protective. Accuracy near 0.7–0.8 is reasonable for a simple, untuned model—good enough for signal, not clinical prediction.
- **Academic impact rates by usage quartile:** The rate table shows the share of students reporting “Yes” to academic impact within each usage quartile. A monotonic rise across quartiles indicates that academic concerns intensify with heavier usage, providing an intuitive complement to the addiction model without another layer of modeling complexity.
- **Actionable summary:** Combine both pieces: higher usage links to higher addiction odds and a larger fraction reporting academic impact. Sleep and mental health still matter, underscoring the intertwined nature of time use and well-being.


# Sensitivity & Robustness

```{r}
# Example:alternate cutpoint
df_alt <- df %>% mutate(high_addiction_alt = factor(if_else(addicted_score >= 8, "High","NotHigh")))
table(df$high_addiction, df_alt$high_addiction_alt)
```

------------------------------------------------------------------------

# What this *means*

-   **Sleep tax:** For many students, more screen time is associated with less sleep (nonlinear + subgroup differences).

-   **Profiles exist:** Clusters capture meaningful combinations of usage/well-being; some groups show elevated conflicts and academic concerns.\

-   **Risk signals:** Predictive models highlight key drivers (usage intensity, sleep, conflicts, platform, demographics).\

-   **Limitations:** Self-report, cross-sectional, measurement error, cross-country comparability.
