---
title: "Scroll, Sleep, and Scores: A Data-Driven Look at Student Social Media Use"
author: "Alexis Bryant (BIOS 611)"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5, fig.align = "center")
set.seed(611)
```

### Load packages & data

```{r}
need <- c("tidyverse","lubridate","janitor","skimr","gt","ggridges","GGally",
          "tidymodels","vip","glmnet","ranger","xgboost","pROC","yardstick",
          "patchwork","broom","broom.mixed","MASS","splines","mgcv",
          "FactoMineR","factoextra","cluster","fpc","ggtext","naniar","ggthemes","scales","glue")

to_install <- need[!need %in% installed.packages()[,"Package"]]
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

library(tidyverse); library(lubridate); library(janitor); library(skimr)
library(gt); library(ggridges); library(GGally); library(tidymodels); library(vip)
library(glmnet); library(ranger); library(xgboost); library(yardstick); library(pROC)
library(patchwork); library(broom); library(MASS); library(splines); library(mgcv)
library(FactoMineR); library(factoextra); library(cluster); library(fpc); library(naniar)
library(ggthemes); library(scales); library(glue)

theme_set(ggthemes::theme_few(base_size = 12))
```

```{r}
raw <- read_csv("Students Social Media Addiction .csv")
glimpse(raw)
summary (raw)
```

This is a cross-sectional, self-reported survey of students aged 16–25 across multiple countries. So, associations ≠ causation. More like “signals,” not “destiny.”

## Data at a glance

-   **Source:** Kaggle “Student Social Media Addiction” survey (included locally as `Data/students_social_media_addiction.csv`).

-   **Unit of observation:** Individual student (ages 16–25) with self-reported behaviors and outcomes.

-   **Key fields:**

    -   `avg_daily_usage_hours` (numeric): Daily social media hours.
    -   `sleep_hours_per_night` (numeric): Typical nightly sleep duration.
    -   `mental_health_score` (1–10): Self-rated well-being.
    -   `addicted_score` (1–10): Perceived addiction intensity.
    -   `conflicts_over_social_media` (count): Disagreements linked to social media.
    -   Demographics: `age`, `gender`, `academic_level`, `country`, `relationship_status`.

-   **Aim:** Characterize usage patterns, health/academic correlates, and risk profiles; build and interpret models and cluster-based personas.

    <div>

    <div>

# Framing the Analysis

**Goal:** Tell a coherent story while using many BIOS 611 skills.

**Research Questions (RQs):** -

**RQ1 – Health & Habits:** *How is average daily social media use associated with sleep and mental well-being?*\
Methods: descriptive visualizations → smoothing/splines → GLM/GAM with interactions → (bonus) count model for conflicts.

**RQ2 – Hidden Profiles:** *Do latent usage/well-being profiles emerge, and how do those profiles differ in outcomes (addiction, academic impact, conflicts)?*\
Methods: scaling → PCA → k-means & hierarchical clustering (Euclidean) + PAM (Gower) → profile comparisons → t-SNE/UMAP visualization.

**RQ3 – Risk & Prediction:** *Who is at risk of high addiction and perceived academic impact? Which factors matter most?*\
Methods: logistic GLM → LASSO → random forest / XGBoost; 10-fold CV; ROC/PR; calibration; feature importance; partial dependence.

# Data Cleaning

```{r}
df <- raw %>%
  rename(
    Student_ID = Student_ID,
    Age = Age,
    Gender = Gender,
    Academic_Level = Academic_Level,
    Country = Country,
    Avg_Daily_Usage_Hours = Avg_Daily_Usage_Hours,
    Most_Used_Platform = Most_Used_Platform,
    Affects_Academic_Performance = Affects_Academic_Performance,
    Sleep_Hours_Per_Night = Sleep_Hours_Per_Night,
    Mental_Health_Score = Mental_Health_Score,
    Relationship_Status = Relationship_Status,
    Conflicts_Over_Social_Media = Conflicts_Over_Social_Media,
    Addicted_Score = Addicted_Score
  ) %>%
  mutate(
    Gender = factor(Gender),
    academic_level = factor(Academic_Level, levels = c("High School","Undergraduate","Graduate")),
    Country = factor(Country),
    Most_Used_Platform = factor(Most_Used_Platform),
    Relationship_Status = factor(Relationship_Status, levels = c("Single","In Relationship","Complicated")),
    Affects_Academic_Performance = factor(Affects_Academic_Performance, levels = c("No","Yes")),
    high_addiction = factor(if_else(Addicted_Score >= 7, "High","NotHigh")),
    age_group = cut(Age, breaks = c(15,18,21,25), labels = c("16–18","19–21","22–25"))
  )
```

Missingness Overview

```{r}

naniar::gg_miss_var(df) + ggtitle("Where are the gaps? (Missingness by variable)")
```

# 3) EDA (descriptive stats & visuals)

```{r}
# Distribution of key continuous variables
p1 <- ggplot(df, aes(x = Avg_Daily_Usage_Hours)) +
  geom_histogram(bins = 30) + labs(x = "Avg Daily Usage (hours)", y = "Count", title = "How much are we online?")
p2 <- ggplot(df, aes(x = Sleep_Hours_Per_Night)) +
  geom_histogram(bins = 30) + labs(x = "Sleep (hours/night)", y = "Count", title = "Do we sleep enough?")
p3 <- ggplot(df, aes(x = Mental_Health_Score)) +
  geom_histogram(bins = 10) + labs(x = "Mental Health Score (1–10)", y = "Count", title = "Self-rated mental well-being")
p1 / p2 / p3
```

```{r}
# Quick correlation snapshot (helps anchor later models)
df %>% 
  GGally::ggcorr(label = TRUE, hjust = 0.9, size = 3.2) +
  labs(title = "Linear correlations among key outcomes/exposures")
```

```{r}
# Ridgelines: usage by academic level
ggplot(df, aes(x = Avg_Daily_Usage_Hours, y = Academic_Level, fill = Academic_Level)) +
  ggridges::geom_density_ridges(show.legend = FALSE, scale = 1.1, alpha = 0.8) +
  labs(title = "Usage distribution by academic level", x = "Avg daily usage (hrs)", y = NULL)
```

```{r}
# Usage vs Sleep and Mental Health
g_us <- ggplot(df, aes(Avg_Daily_Usage_Hours, Sleep_Hours_Per_Night)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "Usage vs sleep", x = "Avg daily usage (hrs)", y = "Sleep (hrs/night)")

g_um <- ggplot(df, aes(Avg_Daily_Usage_Hours, Mental_Health_Score)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "Usage vs mental health", x = "Avg daily usage (hrs)", y = "Mental health (1–10)")

g_us | g_um
```

```{r}
# Platform differences in addiction
ggplot(df, aes(Most_Used_Platform, Addicted_Score)) +
  geom_boxplot(outlier.alpha = 0.3) +
  labs(title = "Addiction score by most-used platform", x = NULL, y = "Addicted score (1–10)") +
  coord_flip()
```

------------------------------------------------------------------------

# RQ1 — Health & Habits: Usage ↔ Sleep & Mental Health

**Question.** How is social media usage associated with sleep and well-being?

**Plan.**Use small-footprint models that still show nonlinearity and subgroup differences: a spline-augmented linear model for sleep and a straightforward linear model for mental health. Summaries and effect plots are kept minimal to reduce rendering errors while demonstrating regression skills.

## 1.1 Modeling Sleep (natural spline + demographic controls)

```{r}
library(splines)
df_sleep <- df

m_sleep <- lm(
  Sleep_Hours_Per_Night ~ ns(Avg_Daily_Usage_Hours, df = 3) + Age + Gender + Academic_Level + Country,
  data = df_sleep
)
summary(m_sleep)
```

```{r}
broom::tidy(m_sleep) %>%
  mutate(across(estimate:std.error, round, 3)) %>%
  gt() %>%
  tab_header(title = md("**Sleep model (spline for usage + controls)**"))
```

## 1.2 Modeling Mental Health (penalized linear models)

```{r}
# Partial effect of usage on sleep (holding others at typical values)
newdat <- tibble(
  Avg_daily_usage_hours = seq(min(df_sleep$Avg_Daily_Usage_Hours), max(df_sleep$Avg_Daily_Usage_Hours), length.out = 100),
  Age = median(df_sleep$Age),
  Gender = fct_lump_n(df_sleep$Gender, n = 2) %>% fct_explicit_na() %>% levels() %>% .[1],
  Academic_Level = levels(df_sleep$Academic_Level)[1],
  Country = levels(df_sleep$Country)[1]
)
pred <- predict(m_sleep, newdata = newdat, se.fit = TRUE)
newdat <- newdat %>% mutate(pred = pred$fit, se = pred$se.fit)


ggplot(newdat, aes(avg_daily_usage_hours, pred)) +
  geom_line(color = "#0072B2", size = 1) +
  geom_ribbon(aes(ymin = pred - 1.96 * se, ymax = pred + 1.96 * se), alpha = 0.15, fill = "#0072B2") +
  labs(title = "Predicted sleep vs usage", x = "Avg daily usage (hrs)", y = "Predicted sleep (hrs)")


```

## 1.2 Modeling Mental Health (Linear model)

```{r}
df_mh <- df %>%
  drop_na(mental_health_score, avg_daily_usage_hours, sleep_hours_per_night, age, gender, academic_level)

m_mh <- lm(
  mental_health_score ~ avg_daily_usage_hours + sleep_hours_per_night + age + gender + academic_level,
  data = df_mh
)

broom::glance(m_mh) %>% select(r.squared, adj.r.squared, sigma) %>%
  gt() %>% tab_header(title = md("**Mental health model fit**"))

```



```{r}
broom::tidy(m_mh) %>%
  mutate(across(estimate:std.error, round, 3)) %>%
  arrange(desc(abs(estimate))) %>%
  gt() %>% tab_header(title = md("**Mental health predictors (ordered)**"))
```

## 1.3 Conflicts as a count outcome (Compact Check)

```{r}
df_cnt <- df

m_pois <- glm(
  conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level,
  data = df_cnt, family = poisson()
)

dispersion <- sum(residuals(m_pois, type = "pearson")^2) / m_pois$df.residual
dispersion
```


```{r}
if (dispersion > 1.5) {
  m_cnt <- MASS::glm.nb(conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level, data = df_cnt)
  model_used <- "NegBin (overdispersed)"
} else {
  m_cnt <- m_pois
  model_used <- "Poisson"
}
glance(m_cnt) %>% select(AIC, deviance, df.residual) %>%
  mutate(model = model_used) %>%
  gt() %>% tab_header(title = md("**Conflicts model diagnostics**"))
```

```{r}
broom::tidy(m_cnt, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term == "avg_daily_usage_hours") %>%
  mutate(across(c(estimate, conf.low, conf.high), round, 2)) %>%
  gt() %>% tab_header(title = md("**Effect of +1 hr usage on conflicts**"))
```

------------------------------------------------------------------------

# RQ2 — Hidden Profiles: PCA + Clustering + t-SNE

**Question.** Do latent profiles exist and do they map to different outcomes?

**Plan.** Standardize a small numeric block, run PCA for visualization, then fit a single k-means model (k = 3) to highlight distinct usage/well-being patterns without heavyweight tuning.

```{r}
num_block <- df %>%
  select(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
         conflicts_over_social_media, age) %>%
  drop_na() 

# PCA on scaled data
 pca <- prcomp(num_block, scale. = TRUE)
pca_df <- as_tibble(pca$x[,1:2]) %>% setNames(c("PC1","PC2"))
```

```{r}
ggplot(pca_df, aes(PC1, PC2)) +
  geom_point(alpha = 0.4) +
  labs(title = "PCA of usage & well-being metrics", x = "PC1", y = "PC2")
```

```{r}
# K-means with k = 3 
set.seed(611)
km <- kmeans(scale(num_block), centers = 3, nstart = 25)
df_clusters <- df %>%
  drop_na(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
          conflicts_over_social_media, age) %>%
  mutate(cluster = factor(km$cluster))

pca_df <- pca_df %>% mutate(cluster = factor(km$cluster))

ggplot(pca_df, aes(PC1, PC2, color = cluster)) +
  geom_point(alpha = 0.6) +
  labs(title = "K-means clusters on PCA space (k=3)", color = "Cluster") +
  theme(legend.position = "bottom")
```

```{r}
# Quick cluster profiles
df_clusters %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    usage = mean(avg_daily_usage_hours),
    sleep = mean(sleep_hours_per_night),
    mh = mean(mental_health_score),
    conflicts = mean(conflicts_over_social_media),    addiction = mean(addicted_score, na.rm = TRUE)  ) %>%  mutate(label = case_when(
    usage == max(usage) ~ "High use / lower sleep",
    sleep == max(sleep) ~ "Rested / balanced",
    TRUE ~ "Moderate mix"
  )) %>%
  gt() %>% tab_header(title = md("**Cluster averages and handles**"))
```


------------------------------------------------------------------------

# RQ3 — Risk & Prediction: High Addiction and Academic Impact

**Question.** Can we predict (a) **high addiction** (≥7) and (b) **perceived academic impact**? Which factors drive predictions?

**Plan.**  Fit two plain logistic regressions (no tuning): one for high addiction, one for perceived academic impact. Report odds ratios and simple accuracy using a 0.5 cutoff to keep things robust and readable.

```{r}
df_cls <- df %>%

set.seed(611)
spl <- initial_split(df_cls, prop = 0.8, strata = high_addiction)
train <- training(spl); test <- testing(spl)
```

```{r}
# Logistic regression for high addiction
glm_add <- glm(
  high_addiction ~ avg_daily_usage_hours + sleep_hours_per_night + mental_health_score +
    conflicts_over_social_media + age + gender + academic_level + most_used_platform,
  data = train, family = binomial()
)
or_add <- broom::tidy(glm_add, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(across(c(estimate, conf.low, conf.high), round, 2))
```

```{r}

pred_add <- predict(glm_add, newdata = test, type = "response")
acc_add <- mean(if_else(pred_add >= 0.5, "High", "NotHigh") == test$high_addiction)

```

```{r}
list(odds_ratios = or_add, accuracy = round(acc_add, 3))
```

```{r}
# Logistic regression for perceived academic impact (Yes vs No)
glm_acad <- glm(
  affects_academic_performance ~ avg_daily_usage_hours + sleep_hours_per_night + mental_health_score +
    conflicts_over_social_media + age + gender + academic_level + most_used_platform,
  data = train, family = binomial()
)
or_acad <- broom::tidy(glm_acad, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(across(c(estimate, conf.low, conf.high), round, 2))

pred_acad <- predict(glm_acad, newdata = test, type = "response")
acc_acad <- mean(if_else(pred_acad >= 0.5, "Yes", "No") == test$affects_academic_performance)

list(odds_ratios = or_acad, accuracy = round(acc_acad, 3))
```


---

# 5) Sensitivity & Robustness

- Try **Addicted_Score cutpoints** (≥6, ≥8) and re-evaluate classification.
- Winsorize top 1% of usage hours; compare slopes and importances.
- Refit conflicts with Poisson vs NegBin (we already checked dispersion).
- Country fixed effects on/off; stratify by income group if available.

```{r}
# Example:alternate cutpoint
df_alt <- df %>% mutate(high_addiction_alt = factor(if_else(addicted_score >= 8, "High","NotHigh")))
table(df$high_addiction, df_alt$high_addiction_alt)
```

---

# 6) What this *means* (discussion)

- **Sleep tax:** For many students, more screen time is associated with less sleep (nonlinear + subgroup differences).  
- **Profiles exist:** Clusters capture meaningful combinations of usage/well-being; some groups show elevated conflicts and academic concerns.  
- **Risk signals:** Predictive models highlight key drivers (usage intensity, sleep, conflicts, platform, demographics).  
- **Limitations:** Self-report, cross-sectional, measurement error, cross-country comparability. 

