---
title: "Scroll, Sleep, and Scores: A Data-Driven Look at Student Social Media Use"
author: "Alexis Bryant (BIOS 611)"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
    theme: flatly
    df_print: paged
params:
  data_path: "Data/students_social_media_addiction.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5, fig.align = "center")
set.seed(611)
```

# 0) Load packages & data

```{r}
# Helper to install if missing
need <- c("tidyverse","lubridate","janitor","skimr","gt","ggridges","GGally",
          "tidymodels","vip","glmnet","ranger","xgboost","pROC","yardstick",
          "patchwork","broom","broom.mixed","MASS","splines","mgcv",
          "FactoMineR","factoextra","cluster","fpc","ggtext","naniar","ggthemes","scales","glue")

to_install <- need[!need %in% installed.packages()[,"Package"]]
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

library(tidyverse); library(lubridate); library(janitor); library(skimr)
library(gt); library(ggridges); library(GGally); library(tidymodels); library(vip)
library(glmnet); library(ranger); library(xgboost); library(yardstick); library(pROC)
library(patchwork); library(broom); library(MASS); library(splines); library(mgcv)
library(FactoMineR); library(factoextra); library(cluster); library(fpc); library(naniar)
library(ggthemes); library(scales); library(glue)

theme_set(ggthemes::theme_few(base_size = 12))
```

```{r}
# Read data
raw <- readr::read_csv(params$data_path) %>% clean_names()
glimpse(raw)
```

> **Vibe check:** This is a cross-sectional, self-reported survey of students aged 16–25 across multiple countries. We’re going to keep it real: associations ≠ causation. Think “signals,” not “destiny.”

## Dataset at a glance

- **Source:** Kaggle “Student Social Media Addiction” survey (included locally as `Data/students_social_media_addiction.csv`).
- **Unit of observation:** Individual student (ages 16–25) with self-reported behaviors and outcomes.
- **Key fields:**
  - `avg_daily_usage_hours` (numeric): Daily social media hours.
  - `sleep_hours_per_night` (numeric): Typical nightly sleep duration.
  - `mental_health_score` (1–10): Self-rated well-being.
  - `addicted_score` (1–10): Perceived addiction intensity.
  - `conflicts_over_social_media` (count): Disagreements linked to social media.
  - Demographics: `age`, `gender`, `academic_level`, `country`, `relationship_status`.
- **Aim:** Characterize usage patterns, health/academic correlates, and risk profiles; build interpretable models and cluster-based personas.

# 1) Framing the journey

**Goal:** Tell a coherent story while flexing many BIOS 611 skills.

**Research Questions (RQs):**
- **RQ1 – Health & Habits:** *How is average daily social media use associated with sleep and mental well-being across student subgroups?*  
  Methods: descriptive viz → smoothing/splines → GLM/GAM with interactions → (bonus) count model for conflicts.
- **RQ2 – Hidden Profiles:** *Do latent usage/well-being profiles emerge, and how do those profiles differ in outcomes (addiction, academic impact, conflicts)?*  
  Methods: scaling → PCA → k-means & hierarchical clustering (Euclidean) + PAM (Gower) → profile comparisons → t-SNE/UMAP visualization.
- **RQ3 – Risk & Prediction:** *Who is at risk of high addiction and perceived academic impact? Which factors matter most?*  
  Methods: logistic GLM → LASSO → random forest / XGBoost; 10-fold CV; ROC/PR; calibration; feature importance; partial dependence.

> **Tone:** conversational with light humor. E.g., “Each extra hour online might be costing you ~X minutes of sleep — your pillow filed a complaint.”

# 2) Data hygiene

```{r}
df <- raw %>%
  rename(
    student_id = student_id,
    age = age,
    gender = gender,
    academic_level = academic_level,
    country = country,
    avg_daily_usage_hours = avg_daily_usage_hours,
    most_used_platform = most_used_platform,
    affects_academic_performance = affects_academic_performance,
    sleep_hours_per_night = sleep_hours_per_night,
    mental_health_score = mental_health_score,
    relationship_status = relationship_status,
    conflicts_over_social_media = conflicts_over_social_media,
    addicted_score = addicted_score
  ) %>%
  mutate(
    gender = factor(gender),
    academic_level = factor(academic_level, levels = c("High School","Undergraduate","Graduate")),
    country = factor(country),
    most_used_platform = factor(most_used_platform),
    relationship_status = factor(relationship_status, levels = c("Single","In Relationship","Complicated")),
    affects_academic_performance = factor(affects_academic_performance, levels = c("No","Yes")),
    high_addiction = factor(if_else(addicted_score >= 7, "High","NotHigh")),
    age_group = cut(age, breaks = c(15,18,21,25), labels = c("16–18","19–21","22–25"))
  )
```

```{r}
# Missingness overview
naniar::gg_miss_var(df) + ggtitle("Where are the gaps? (Missingness by variable)")
```

```{r}
# Quick skim
skimr::skim(df %>% select(-student_id))
```

> **Quality check:** The skim output is a quick audit for data surprises. If you see heavy missingness in any driver (e.g., sleep hours), note it in the write-up and consider sensitivity checks that drop or impute those rows.

# 3) Warm-up EDA (descriptives & effective visuals)

```{r}
# Distribution of key continuous variables
p1 <- ggplot(df, aes(x = avg_daily_usage_hours)) +
  geom_histogram(bins = 30) + labs(x = "Avg Daily Usage (hours)", y = "Count", title = "How much are we online?")
p2 <- ggplot(df, aes(x = sleep_hours_per_night)) +
  geom_histogram(bins = 30) + labs(x = "Sleep (hours/night)", y = "Count", title = "Do we sleep enough?")
p3 <- ggplot(df, aes(x = mental_health_score)) +
  geom_histogram(bins = 10) + labs(x = "Mental Health Score (1–10)", y = "Count", title = "Self-rated mental well-being")
p1 / p2 / p3
```

```{r}
# Quick correlation snapshot (helps anchor later models)
df %>%
  select(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score, conflicts_over_social_media) %>%
  drop_na() %>%
  GGally::ggcorr(label = TRUE, hjust = 0.9, size = 3.2) +
  labs(title = "Linear correlations among key outcomes/exposures")
```

```{r}
# Ridgelines: usage by academic level
ggplot(df, aes(x = avg_daily_usage_hours, y = academic_level, fill = academic_level)) +
  ggridges::geom_density_ridges(show.legend = FALSE, scale = 1.1, alpha = 0.8) +
  labs(title = "Usage distribution by academic level", x = "Avg daily usage (hrs)", y = NULL)
```

```{r}
# Usage vs Sleep and Mental Health
g_us <- ggplot(df, aes(avg_daily_usage_hours, sleep_hours_per_night)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "More scrolling, less snoozing?", x = "Avg daily usage (hrs)", y = "Sleep (hrs/night)")

g_um <- ggplot(df, aes(avg_daily_usage_hours, mental_health_score)) +
  geom_point(alpha = 0.4) + geom_smooth(se = TRUE, method = "loess") +
  labs(title = "Usage vs mental health", x = "Avg daily usage (hrs)", y = "Mental health (1–10)")

g_us | g_um
```

```{r}
# Platform differences in addiction
ggplot(df, aes(most_used_platform, addicted_score)) +
  geom_boxplot(outlier.alpha = 0.3) +
  labs(title = "Addiction score by most-used platform", x = NULL, y = "Addicted score (1–10)") +
  coord_flip()
```

---

# RQ1 — Health & Habits: Usage ↔ Sleep & Mental Health

**Question.** How is social media usage associated with sleep and well-being, and do effects vary by age group and academic level?

**Plan.** Explore nonlinearity (splines/GAM), interactions, and model conflicts as a count (Poisson vs NegBin).

## 1.1 Modeling Sleep (linear with splines)

```{r}
# Linear model with spline for usage, interactions with academic level
m_sleep <- mgcv::gam(sleep_hours_per_night ~ s(avg_daily_usage_hours, k = 5) +
                       age + gender + academic_level + country +
                       s(avg_daily_usage_hours, by = academic_level, k = 5),
                     data = df, method = "REML")
summary(m_sleep)
```

```{r}
# Partial effect plot for usage by level
newdat <- tidyr::expand_grid(
  avg_daily_usage_hours = seq(min(df$avg_daily_usage_hours, na.rm=TRUE),
                              max(df$avg_daily_usage_hours, na.rm=TRUE), length.out = 100),
  academic_level = levels(df$academic_level),
  age = median(df$age, na.rm=TRUE),
  gender = fct_lump_n(df$gender, n = 2) %>% levels() %>% .[1],
  country = levels(df$country)[1]
)

newdat$pred <- predict(m_sleep, newdata = newdat, type = "response", se.fit = TRUE)$fit
newdat$se <- predict(m_sleep, newdata = newdat, type = "response", se.fit = TRUE)$se.fit

ggplot(newdat, aes(avg_daily_usage_hours, pred, color = academic_level)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = pred - 1.96*se, ymax = pred + 1.96*se, fill = academic_level), alpha = 0.15, color = NA) +
  labs(title = "Predicted sleep vs usage by academic level", x = "Avg daily usage (hrs)", y = "Predicted sleep (hrs)") +
  guides(fill = "none")
```

## 1.2 Modeling Mental Health (penalized linear models)

```{r}
# Prepare recipe and data for penalized regression
df_reg <- df %>%
  select(mental_health_score, avg_daily_usage_hours, sleep_hours_per_night, age, gender,
         academic_level, country, most_used_platform, relationship_status, conflicts_over_social_media) %>%
  drop_na()

set.seed(611)
spl <- initial_split(df_reg, prop = 0.8, strata = mental_health_score)
trn <- training(spl); tst <- testing(spl)

rec_reg <- recipe(mental_health_score ~ ., data = trn) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

spec_lasso <- linear_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet")
wf_lasso <- workflow(rec_reg, spec_lasso)

set.seed(611)
folds <- vfold_cv(trn, v = 10)
grid_lambda <- grid_regular(penalty(range = c(-4, 0)), levels = 30)

res_lasso <- tune_grid(wf_lasso, resamples = folds,
                       grid = grid_lambda,
                       metrics = metric_set(rmse, rsq))

best_lasso <- select_best(res_lasso, "rmse")
final_lasso <- finalize_workflow(wf_lasso, best_lasso) %>% fit(trn)

preds <- predict(final_lasso, tst) %>% bind_cols(tst)
metrics <- yardstick::metrics(preds, truth = mental_health_score, estimate = .pred)

metrics
```

```{r}
# Interpret RMSE/RSQ in plain English
mh_take <- metrics %>%
  select(.metric, .estimate) %>%
  mutate(comment = case_when(
    .metric == "rmse" ~ "Average prediction error (on the original 1–10 scale). Lower = better.",
    .metric == "rsq"  ~ "Variance explained by the model. Values >0.2 are often useful for noisy survey data.",
    TRUE ~ ""
  ))
gt(mh_take) %>% tab_header(title = md("**Mental health model performance (test set)**"))
```

```{r}
# Coefficients (non-zero)
lasso_fit <- final_lasso %>% extract_fit_parsnip()
coef_mat <- as.matrix(coef(lasso_fit$fit, s = lasso_fit$fit$lambdaOpt))
nz <- coef_mat[coef_mat[,1]!=0, , drop = FALSE]
tibble(term = rownames(nz), estimate = as.numeric(nz)) %>% arrange(desc(abs(estimate))) %>% head(20)
```

## 1.3 Conflicts as a count outcome (Poisson vs NegBin)

```{r}
df_cnt <- df %>%
  select(conflicts_over_social_media, avg_daily_usage_hours, age, gender, academic_level,
         country, most_used_platform, relationship_status) %>%
  drop_na()

m_pois <- glm(conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level +
                country + most_used_platform + relationship_status,
              data = df_cnt, family = poisson())

dispersion <- sum(residuals(m_pois, type="pearson")^2) / m_pois$df.residual
dispersion

m_nb <- MASS::glm.nb(conflicts_over_social_media ~ avg_daily_usage_hours + age + gender + academic_level +
                       country + most_used_platform + relationship_status, data = df_cnt)

AIC(m_pois); AIC(m_nb)
broom::tidy(m_nb, exponentiate = TRUE, conf.int = TRUE) %>%
  arrange(desc(estimate)) %>% head(12) %>%
  gt() %>% tab_header(title = md("**Rate Ratios for Conflicts (NegBin)**"))
```

```{r}
# Narrative summary for conflicts
rr_usage <- broom::tidy(m_nb, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term == "avg_daily_usage_hours") %>%
  transmute(
    message = glue::glue("Each +1 hour of daily use is associated with a {round(estimate,2)}x higher expected conflict count (95% CI {round(conf.low,2)}–{round(conf.high,2)}).")
  )
rr_usage$message
```

> **Interpretation game plan:** Translate slopes into real terms (e.g., “+1 hr usage ≈ −0.18 hrs sleep”), call out heterogeneity by level, and mention overdispersion choice.

---

# RQ2 — Hidden Profiles: PCA + Clustering + t-SNE

**Question.** Do latent profiles exist and do they map to different outcomes?

**Plan.** Standardize numeric block → PCA → choose k via silhouette/gap → cluster (k-means, Ward, PAM/Gower) → compare profiles on outcomes.

```{r}
num_block <- df %>%
  select(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
         conflicts_over_social_media, age) %>%
  drop_na() %>%
  scale() %>%
  as.data.frame()

# PCA
pca <- FactoMineR::PCA(num_block, graph = FALSE)
factoextra::fviz_eig(pca)
factoextra::fviz_pca_biplot(pca, repel = TRUE, col.var = "black")
```

```{r}
# k-means on PC scores
pc_scores <- as_tibble(pca$ind$coord)
sil_vec <- map_dbl(2:8, ~{
  km <- kmeans(pc_scores, centers = .x, nstart = 25)
  mean(cluster::silhouette(km$cluster, dist(pc_scores))[,3])
})
best_k <- which.max(sil_vec) + 1
best_k
```

```{r}
# Compact cluster diagnostics and plain-language handles
km_summary <- df_clusters %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    usage = mean(avg_daily_usage_hours),
    sleep = mean(sleep_hours_per_night),
    mh = mean(mental_health_score),
    conflicts = mean(conflicts_over_social_media),
    addiction = mean(addicted_score, na.rm=TRUE),
    acad_yes = mean(affects_academic_performance == "Yes", na.rm=TRUE)
  ) %>%
  mutate(label = case_when(
    usage >= max(usage) & sleep <= min(sleep) ~ "High use / low sleep",
    usage <= min(usage) & mh >= max(mh) ~ "Low use / thriving",
    TRUE ~ "Moderate mix"
  ))
km_summary %>% gt() %>% tab_header(title = md("**Cluster profiles (name + averages)**"))
```

```{r}
set.seed(611)
km <- kmeans(pc_scores, centers = best_k, nstart = 50)
df_clusters <- df %>% drop_na(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
                              conflicts_over_social_media, age) %>%
  mutate(cluster = factor(km$cluster))

# Profile plot
df_clusters %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    usage = mean(avg_daily_usage_hours),
    sleep = mean(sleep_hours_per_night),
    mh = mean(mental_health_score),
    conflicts = mean(conflicts_over_social_media),
    addiction = mean(addicted_score, na.rm=TRUE),
    acad_yes = mean(affects_academic_performance == "Yes", na.rm=TRUE)
  ) %>%
  pivot_longer(-cluster, names_to = "metric", values_to = "value") %>%
  ggplot(aes(metric, value, fill = cluster)) + geom_col(position = "dodge") +
  coord_flip() + labs(title = "Cluster profiles across key metrics", x = NULL, y = NULL)
```

```{r}
# PAM with Gower on mixed data (sensitivity)
mix <- df %>%
  select(avg_daily_usage_hours, sleep_hours_per_night, mental_health_score,
         conflicts_over_social_media, age, gender, academic_level, most_used_platform, relationship_status) %>%
  drop_na()

gower <- cluster::daisy(mix, metric = "gower")
pam_fit <- cluster::pam(gower, k = best_k)
silpam <- summary(silhouette(pam_fit))
silpam$avg.width
```

```{r}
# t-SNE visualization (on PC scores) colored by cluster
set.seed(611)
# Using Rtsne if available; otherwise UMAP via umap
if (!requireNamespace("Rtsne", quietly = TRUE)) install.packages("Rtsne", repos = "https://cloud.r-project.org")
library(Rtsne)
ts <- Rtsne(as.matrix(pc_scores), perplexity = 30, check_duplicates = FALSE)
ts_df <- as_tibble(ts$Y) %>% setNames(c("x","y")) %>% mutate(cluster = factor(km$cluster))

ggplot(ts_df, aes(x, y, color = cluster)) + geom_point(alpha = 0.7) +
  labs(title = "t-SNE view of latent profiles", x = NULL, y = NULL) +
  theme(legend.position = "bottom")
```

> **Interpretation:** Name the clusters in plain English (e.g., “High-Use/Low-Sleep”, “Moderate/Stable”, “Low-Use/High-Wellbeing”). Compare outcomes and academic impact across clusters.

---

# RQ3 — Risk & Prediction: High Addiction and Academic Impact

**Question.** Can we predict (a) **high addiction** (≥7) and (b) **perceived academic impact**? Which factors drive predictions?

**Plan.** Logistic GLM → LASSO → Random Forest (and optional XGBoost). 10-fold CV, held-out test set, ROC/PR, calibration. Explain via permutation importance and partial dependence.

```{r}
df_cls <- df %>%
  select(high_addiction, affects_academic_performance, avg_daily_usage_hours,
         sleep_hours_per_night, mental_health_score, age, gender, academic_level,
         country, most_used_platform, relationship_status, conflicts_over_social_media) %>%
  drop_na()

set.seed(611)
split <- initial_split(df_cls, prop = 0.8, strata = high_addiction)
train <- training(split); test <- testing(split)

set.seed(611)
folds <- vfold_cv(train, v = 10, strata = high_addiction)

rec_cls <- recipe(high_addiction ~ ., data = train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# Baseline GLM
glm_spec <- logistic_reg() %>% set_engine("glm")
glm_wf <- workflow(rec_cls, glm_spec)

# LASSO
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet")
lasso_wf <- workflow(rec_cls, lasso_spec)

# Random Forest
rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_engine("ranger", importance = "permutation") %>% set_mode("classification")
rf_wf <- workflow(rec_cls, rf_spec)
```

```{r}
# Tuning
set.seed(611)
glm_res   <- fit_resamples(glm_wf, folds, metrics = metric_set(roc_auc, pr_auc, brier_class))

set.seed(611)
lasso_res <- tune_grid(lasso_wf, folds,
                       grid = grid_regular(penalty(range = c(-4,0)), levels = 30),
                       metrics = metric_set(roc_auc, pr_auc, brier_class))

set.seed(611)
rf_res <- tune_grid(rf_wf, folds,
                    grid = grid_latin_hypercube(mtry(range = c(3, 20)),
                                                min_n(range = c(2, 20)), size = 25),
                    metrics = metric_set(roc_auc, pr_auc, brier_class))

collect_metrics(glm_res) %>% filter(.metric %in% c("roc_auc","pr_auc","brier_class")) %>% mutate(model="GLM") %>% arrange(desc(mean)) %>% head(6)

best_lasso <- select_best(lasso_res, "roc_auc")
best_rf <- select_best(rf_res, "roc_auc")

final_lasso <- finalize_workflow(lasso_wf, best_lasso) %>% fit(train)
final_rf    <- finalize_workflow(rf_wf, best_rf) %>% fit(train)
```

```{r}
# Test performance
eval_model <- function(fitted, data){
  p <- predict(fitted, data, type = "prob") %>% bind_cols(data %>% select(high_addiction))
  list(
    roc = roc_auc(p, truth = high_addiction, .pred_High),
    pr  = pr_auc(p, truth = high_addiction, .pred_High),
    brier = brier_class(p, truth = high_addiction, .pred_High)
  )
}

m_glm   <- glm_wf %>% fit(train)
perf_glm   <- eval_model(m_glm, test)
perf_lasso <- eval_model(final_lasso, test)
perf_rf    <- eval_model(final_rf, test)

perf_glm; perf_lasso; perf_rf
```

```{r}
# Side-by-side model comparison
model_cards <- bind_rows(
  bind_cols(model = "Baseline GLM", tibble::enframe(perf_glm) %>% unnest_wider(value)),
  bind_cols(model = "LASSO", tibble::enframe(perf_lasso) %>% unnest_wider(value)),
  bind_cols(model = "Random Forest", tibble::enframe(perf_rf) %>% unnest_wider(value))
) %>%
  select(model, .metric, .estimate) %>%
  arrange(desc(.metric), desc(.estimate))

model_cards %>% gt() %>% tab_header(title = md("**High-addiction prediction: metric dashboard**"))
```

```{r}
# Calibration plot for RF
rf_preds <- predict(final_rf, test, type = "prob") %>% bind_cols(test %>% select(high_addiction))
cal_df <- rf_preds %>%
  mutate(bin = ntile(.pred_High, 10)) %>%
  group_by(bin) %>%
  summarise(pred = mean(.pred_High), obs = mean(high_addiction == "High"))

ggplot(cal_df, aes(pred, obs)) +
  geom_point(size = 2) + geom_abline(linetype = 2) +
  labs(title = "Calibration: Random Forest for High Addiction", x = "Predicted", y = "Observed")
```

```{r}
# Feature importance + partial dependence
rf_fit <- final_rf %>% extract_fit_parsnip()
vip::vip(rf_fit, num_features = 12)

# Partial dependence for usage and sleep
library(pdp)
p1 <- pdp::partial(rf_fit, pred.var = "avg_daily_usage_hours", train = bake(prep(rec_cls), new_data = train), prob = TRUE) %>% autoplot()
p2 <- pdp::partial(rf_fit, pred.var = "sleep_hours_per_night", train = bake(prep(rec_cls), new_data = train), prob = TRUE) %>% autoplot()
p1 / p2
```

```{r}
# (Optional) Predict perceived academic impact as a second target
df_acad <- df_cls %>%
  mutate(affects = affects_academic_performance) %>%
  select(-high_addiction) %>%
  drop_na()

set.seed(611)
split2 <- initial_split(df_acad, prop = 0.8, strata = affects)
train2 <- training(split2); test2 <- testing(split2)
folds2 <- vfold_cv(train2, v = 10, strata = affects)

rec2 <- recipe(affects ~ ., data = train2) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

rf2 <- rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_engine("ranger", importance = "permutation") %>% set_mode("classification")

rf2_wf <- workflow(rec2, rf2)

set.seed(611)
rf2_res <- tune_grid(rf2_wf, folds2,
                     grid = grid_latin_hypercube(mtry(range = c(3, 20)),
                                                 min_n(range = c(2, 20)), size = 25),
                     metrics = metric_set(roc_auc, pr_auc, brier_class))

best2 <- select_best(rf2_res, "roc_auc")
final_rf2 <- finalize_workflow(rf2_wf, best2) %>% fit(train2)

p2_test <- predict(final_rf2, test2, type = "prob") %>% bind_cols(test2 %>% select(affects))
list(
  roc = roc_auc(p2_test, truth = affects, .pred_Yes),
  pr = pr_auc(p2_test, truth = affects, .pred_Yes),
  brier = brier_class(p2_test, truth = affects, .pred_Yes)
)
```

---

# 5) Sensitivity & Robustness

- Try **Addicted_Score cutpoints** (≥6, ≥8) and re-evaluate classification.
- Winsorize top 1% of usage hours; compare slopes and importances.
- Refit conflicts with Poisson vs NegBin (we already checked dispersion).
- Country fixed effects on/off; stratify by income group if available.

```{r}
# Example:alternate cutpoint
df_alt <- df %>% mutate(high_addiction_alt = factor(if_else(addicted_score >= 8, "High","NotHigh")))
table(df$high_addiction, df_alt$high_addiction_alt)
```

---

# 6) What this *means* (discussion)

- **Sleep tax:** For many students, more screen time is associated with less sleep (nonlinear + subgroup differences).  
- **Profiles exist:** Clusters capture meaningful combinations of usage/well-being; some groups show elevated conflicts and academic concerns.  
- **Risk signals:** Predictive models highlight key drivers (usage intensity, sleep, conflicts, platform, demographics).  
- **Limitations:** Self-report, cross-sectional, measurement error, cross-country comparability.  
- **Next:** Longitudinal follow-up, objective app logs, causal designs.

---

# Appendix

- Data dictionary (from Kaggle description)
- Session info
- Full model specs / tuning results

```{r}
sessionInfo()
```